---
title: "Cell communication_NOX"
output: html_notebook
---

```{r}
library(Seurat)
library(nichenetr)
library(pheatmap)
library(tidyverse)
library(RColorBrewer)
library(DESeq2)
library(muscat)
```

# Load the data
```{r}
memory.limit(size=90000)
seurat <- readRDS("C:/Users/.../seurat_NOX.rds")
result_table <- read.csv("C:/Users/.../DSA_lung_NOX.csv", stringsAsFactors = F)
```


# Prepare NicheNet prior model
```{r}
ligand_target_matrix <- readRDS(url("https://zenodo.org/record/3260758/files/ligand_target_matrix.rds"))
ligand_target_matrix[1:5,1:5] # target genes in rows, ligands in columns
```
```{r}
lr_network <- readRDS(url("https://zenodo.org/record/3260758/files/lr_network.rds"))
head(lr_network)
```
```{r}
weighted_networks <- readRDS(url("https://zenodo.org/record/3260758/files/weighted_networks.rds"))
weighted_networks_lr = weighted_networks$lr_sig %>% inner_join(lr_network %>% distinct(from,to), by = c("from","to"))
```

Convert symbols of the model from human to mouse
```{r}
lr_network = lr_network %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
colnames(ligand_target_matrix) = ligand_target_matrix %>% colnames() %>% convert_human_to_mouse_symbols()
rownames(ligand_target_matrix ) = ligand_target_matrix %>% rownames() %>% convert_human_to_mouse_symbols()
ligand_target_matrix = ligand_target_matrix %>% .[!is.na(rownames(ligand_target_matrix)), !is.na(colnames(ligand_target_matrix))]
weighted_networks_lr = weighted_networks_lr %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
```

```{r}
seurat$celltype <- seurat$CellType_II
```



# Run NicheNet
```{r}
clusters <- result_table %>%
  filter(p_adj.loc <= 0.05 &
           abs(logFC) >= 0.5) %>%
  filter(P3.frq >= 0.1 | P14.frq >= 0.1) %>%
  select(cluster_id) %>%
  group_by(cluster_id) %>%
  summarise(nGene = n())
clusters
```

```{r}
receiver_clusters <- filter(clusters, nGene >= 200) %>%
  pull(cluster_id)
receiver_clusters
```

```{r}
getFrequency <- function(cluster){
  cluster_cells <- colnames(seurat)[seurat$CellType_II == cluster]
  
  subset_mat <- as.matrix(seurat[["RNA"]]@data[, cluster_cells])
  subset_frq <- rowSums(subset_mat > 0) / ncol(subset_mat)
  genes_keep <- rownames(subset_mat)[subset_frq >= 0.1]
  return(genes_keep)
}
```

```{r}
runNicheNet <- function(receiver){
  print(paste0("Performing analysis on cluster: ", receiver))
  #define receiver
  expressed_genes_receiver <- getFrequency(receiver)
  background_expressed_genes <- expressed_genes_receiver %>% .[. %in% rownames(ligand_target_matrix)]
  
  #define senders
  sender_celltypes <- clusters$cluster_id[which(clusters$cluster_id %in% result_table$cluster_id)]
  list_expressed_genes_sender <- sender_celltypes %>% unique() %>% lapply(getFrequency) 
  expressed_genes_sender <- list_expressed_genes_sender %>% unlist() %>% unique()
  
  #get P14 genes of interest
  geneset_oi <- filter(result_table, p_adj.loc <= 0.05 & 
                       abs(logFC) >= 0.5 &
                       cluster_id == receiver)
  geneset_oi <- filter(geneset_oi, P3.frq >= 0.1 | P14.frq >= 0.1) %>% pull(gene)
  
  #define potential ligands for genes
  ligands <- lr_network %>% pull(from) %>% unique()
  receptors <- lr_network %>% pull(to) %>% unique()
  expressed_ligands <- intersect(ligands,expressed_genes_sender)
  expressed_receptors <- intersect(receptors,expressed_genes_receiver)
  potential_ligands <- lr_network %>% 
    filter(from %in% expressed_ligands & to %in% expressed_receptors) %>% 
    pull(from) %>% 
    unique()
  
  #perform nichenet ligand activity analysis: rank potential ligands based on the presence
  #of their target genes in the gene set of interest (compared to the background set of genes)
  ligand_activities <- predict_ligand_activities(geneset = geneset_oi, 
                                              background_expressed_genes = background_expressed_genes, 
                                              ligand_target_matrix = ligand_target_matrix, 
                                              potential_ligands = potential_ligands)
  ligand_activities <- ligand_activities %>% arrange(-pearson) %>% mutate(rank = rank(dplyr::desc(pearson)))
  ligand_activities$receiver_cluster <- receiver
  
  return(ligand_activities)
}
```


```{r}
Sys.time()
ligand_activities <- lapply(receiver_clusters, runNicheNet) 
Sys.time()
```

```{r}
ligand_activities <- do.call("rbind", ligand_activities)
```

```{r}
saveRDS(ligand_activities, file="C:/Users/.../ligand_activities_NOX.rds")
```

```{r}
ligand_activities <- readRDS("C:/Users/...ligand_activities_NOX.rds")
```


```{r}
pearson_scores <- ggplot(ligand_activities, aes(x=receiver_cluster, y=pearson)) + 
  geom_jitter(size=0.5) + 
  theme_bw() + 
  theme(axis.text.x=element_text(angle=45, hjust=1))
pearson_scores
```

# Summary figures for each subset
```{r}
subset_dict <- data.frame(subset = seurat$cluster_high_res,
                          CellType_II = seurat$CellType_II)
subset_dict <- unique(subset_dict)
subset_dict <- separate(subset_dict, subset, c("subset", "clusterID"))
subset_dict$clusterID <- NULL
subset_dict$subset <- as.character(subset_dict$subset)
subset_dict$CellType_II <- as.character(subset_dict$CellType_II)
```


# Ligand plot functions
## Ligand activity
```{r}
ligand_auroc_heatmap <- function(subset_name, plot_width, plot_height){
  
  subset_clusters <- subset_dict %>% filter(subset == subset_name) %>% pull(CellType_II)
  
  ligands <- ligand_activities %>%
    filter(receiver_cluster %in% subset_clusters &
           rank <= 10)
  
  ligand_mat <- ligands[,c("test_ligand", "auroc", "receiver_cluster")]
  ligand_mat <- pivot_wider(ligand_mat, names_from="receiver_cluster", values_from="auroc")
  ligand_mat <- as.data.frame(ligand_mat)
  ligand_list <- ligand_mat$test_ligand
  cell_types <- colnames(ligand_mat)[2:ncol(ligand_mat)]
  ligand_mat <- as.matrix(ligand_mat[,2:ncol(ligand_mat)])
  rownames(ligand_mat) <- ligand_list
  colnames(ligand_mat) <- cell_types
  ligand_mat[is.na(ligand_mat)] <- 0
  ligand_mat[ligand_mat > 0] <- 1
  
  #Can't do clustered columns if only one cell type
  if(length(unique(ligands$receiver_cluster)) > 1){
    ligand_heatmap <- pheatmap(ligand_mat,
         color = c("whitesmoke", "firebrick"),
         legend = F,
         cluster_rows=T,
         cluster_cols=T,
         show_rownames=T,
         show_colnames=T,
         treeheight_row=0,
         treeheight_col=0,
         clustering_method="ward.D2",
         border_color = "black",
         filename=paste0("C:/Users/.../", 
                         subset_name, "_ligand_activity_NOX.png"),
         width = plot_width,
         height = plot_height)
  } else{
    ligand_heatmap <- pheatmap(ligand_mat,
         color = c("firebrick", "firebrick"),
         breaks = c(0,1),
         legend = F,
         cluster_rows=F,
         cluster_cols=F,
         show_rownames=T,
         show_colnames=T,
         treeheight_row=0,
         treeheight_col=0,
         clustering_method="ward.D2",
         border_color = "black",
         filename=paste0("C:/Users/.../", 
                         subset_name, "_ligand_activity_NOX.png"),
         width = plot_width,
         height = plot_height)
  }
  
  return(ligand_heatmap)
}
```

```{r}
unique(subset_dict$subset)
```

```{r}
ligand_stroma <- ligand_auroc_heatmap("Stroma", 1.8, 7.5)
ligand_myeloid <- ligand_auroc_heatmap("Myeloid", 1.4, 6.0)
ligand_epithelial <- ligand_auroc_heatmap("Epithelial", 1.4, 6.0)
ligand_lymphocytes <- ligand_auroc_heatmap("Lymphocytes", 1.25, 5.8) 
ligand_endothelial <- ligand_auroc_heatmap("Endothelial", 1.4, 6.0)
ligand_mesothelial <- ligand_auroc_heatmap("Mesothelial", 1.1, 4) 
```


```{r}
Idents(seurat) <- "CellType_II"
```

## Average expression
```{r}
getAvgExp <- function(cluster, gene_order){
  cells <- colnames(seurat)[seurat$CellType_II == cluster & seurat$Age == "P14"]
  avg <- rowMeans(as.matrix(seurat[["RNA"]]@data[gene_order,cells]))
  return(avg)
}
```

```{r}
ligand_expression_heatmap <- function(ligand_heatmap, subset_name,
                                      plot_height){
  
  #Get ligand list from heatmap
  gene_order <- ligand_heatmap$gtable$grobs[[3]]$label
  
  clusters <- levels(Idents(seurat))
  ligand_avg <- lapply(clusters, getAvgExp, gene_order=gene_order)
  ligand_avg <- do.call("cbind", ligand_avg)
  colnames(ligand_avg) <- clusters
  
  ligand_exp_heatmap <- pheatmap(ligand_avg,
         color = colorRampPalette(c("whitesmoke", "purple"))(100),
         breaks = seq(0, 2, length.out=101),
         cluster_rows=F,
         cluster_cols=F,
         show_rownames=T,
         show_colnames=T,
         treeheight_row=0,
         treeheight_col=0,
         clustering_method="ward.D2",
         border_color = "black",
         filename=paste0("C:/Users/.../", subset_name, "_ligand_expression_NOX.png"),
         width=7, 
         height=plot_height)
  
}
```

```{r}
ligand_expression_heatmap(ligand_stroma, "Stroma", 7)
ligand_expression_heatmap(ligand_myeloid, "Myeloid", 6.7)
ligand_expression_heatmap(ligand_epithelial, "Epithelial", 6.7)
ligand_expression_heatmap(ligand_endothelial, "Endothelial", 6.7)
ligand_expression_heatmap(ligand_lymphocytes, "Lymphocytes", 5)
ligand_expression_heatmap(ligand_mesothelial, "Mesothelial", 4)
```

## Fold change
```{r}
getFoldChange <- function(cluster, gene_order){
  filtered_table <- filter(result_table, cluster_id == cluster &
                             p_adj.loc <= 0.05)
  
  fc <- filtered_table$logFC
  names(fc) <- filtered_table$gene
  fc <- fc[gene_order]
  names(fc) <- gene_order
  fc[is.na(fc)] <- 0
  
  return(fc)
}
```

```{r}
ligand_fc_heatmap <- function(ligand_heatmap, subset_name, plot_height){
  gene_order <- ligand_heatmap$gtable$grobs[[3]]$label
  clusters <- levels(Idents(seurat))
  
  fold_change <- lapply(clusters, getFoldChange, gene_order = gene_order)
  
  fold_change <- do.call("cbind", fold_change)
  colnames(fold_change) <- clusters
  
  ligand_fc_heatmap <- pheatmap(fold_change,
         color = colorRampPalette(rev(brewer.pal(7, "RdBu")))(100),
         breaks = seq(-2, 2, length.out=101),
         cluster_rows=F,
         cluster_cols=F,
         show_rownames=T,
         show_colnames=T,
         treeheight_row=0,
         treeheight_col=0,
         clustering_method="ward.D2",
         border_color = "black",
         filename=paste0("C:/Users/.../", subset_name, "_ligand_foldchange_NOX.png"),
         width=7, height=plot_height)
}
```

```{r}
ligand_fc_heatmap(ligand_stroma, "Stroma", 7)
ligand_fc_heatmap(ligand_myeloid, "Myeloid", 6.7)
ligand_fc_heatmap(ligand_epithelial, "Epithelial", 6.7)
ligand_fc_heatmap(ligand_endothelial, "Endothelial", 6.7)
ligand_fc_heatmap(ligand_lymphocytes, "Lymphocytes", 5)
ligand_fc_heatmap(ligand_mesothelial, "Mesothelial", 4)
```

# Putative receptors
## Quick check on distribution of ligand-receptor weights in the network
```{r}
hist(weighted_networks_lr$weight, breaks=50)
```

## Plot receptor weights
```{r}
getFrequency <- function(cluster){
  cluster_cells <- colnames(seurat)[Idents(seurat) == cluster]
  receptors <- lr_network %>% pull(to) %>% unique()
  receptors_check <- receptors[which(receptors %in% rownames(seurat))]
  
  subset_mat <- as.matrix(seurat[["RNA"]]@data[receptors_check, cluster_cells])
  subset_frq <- rowSums(subset_mat > 0) / ncol(subset_mat)
  genes_keep <- rownames(subset_mat)[subset_frq >= 0.1]
  return(genes_keep)
}
```

```{r}
receptor_interaction <- function(subset_name, ligand_heatmap, plot_width, plot_height){
  #Get all cell types
  subset_clusters <- subset_dict %>% filter(subset == subset_name) %>% pull(CellType_II)
  
  #Get ordered ligand list from heatmap object
  ligand_order <- ligand_heatmap$gtable$grobs[[3]]$label
  
  #Define target receptors
  receptors <- lr_network %>% pull(to) %>% unique()
  ##Iterate through clusters checking for frequency of expressing cells
  expressed_receptors <- lapply(subset_clusters, getFrequency)
  expressed_receptors <- unique(unlist(expressed_receptors))
  ##Filter network
  lr_network_top = lr_network %>% 
    filter(from %in% ligand_order & to %in% expressed_receptors) %>% 
    distinct(from,to)
  best_upstream_receptors = lr_network_top %>% pull(to) %>% unique()
  
  #Find ligand-receptor network
  lr_network_top_df_large = weighted_networks_lr %>% 
    filter(from %in% ligand_order & to %in% best_upstream_receptors)
  ##Make a wide format
  lr_network_top_df = lr_network_top_df_large %>% spread("from","weight",fill = 0)
  lr_network_top_matrix = lr_network_top_df %>% 
    select(-to) %>% 
    as.matrix() %>% 
    magrittr::set_rownames(lr_network_top_df$to)
  
  #Cluster receptors
  dist_receptors = dist(lr_network_top_matrix, method = "binary")
  hclust_receptors = hclust(dist_receptors, method = "ward.D2")
  order_receptors = hclust_receptors$labels[hclust_receptors$order]
  
  #Test heatmap
  receptor_mat <- t(lr_network_top_matrix[order_receptors, ligand_order])
  
  ligand_receptor_heatmap <- pheatmap(receptor_mat,
         color = colorRampPalette(c("whitesmoke", "darkgreen"))(100),
         breaks = seq(0, 1, length.out=101),
         cluster_rows=F,
         cluster_cols=F,
         show_rownames=T,
         show_colnames=T,
         treeheight_row=0,
         treeheight_col=0,
         clustering_method="ward.D2",
         border_color = "black",
         filename=paste0("C:/Users/.../", subset_name, "_receptor_activity_NOX.png"),
         width=plot_width, 
         height=plot_height)
  return(ligand_receptor_heatmap)
}
```

```{r}
receptor_stroma <- receptor_interaction("Stroma", ligand_stroma, 17, 5.8)
receptor_myeloid <- receptor_interaction("Myeloid", ligand_myeloid, 20, 5)
receptor_endothelial <- receptor_interaction("Endothelial", ligand_endothelial, 12, 4.5)
receptor_epithelial <- receptor_interaction("Epithelial", ligand_epithelial, 17, 5.8)
receptor_lymphocytes <- receptor_interaction("Lymphocytes", ligand_lymphocytes, 12, 4)
receptor_mesothelial <- receptor_interaction("Mesothelial", ligand_mesothelial, 8, 2.5)
```

# Receptor expression in receivers
```{r}
getAvgExp <- function(cluster, gene_order){
  cells <- colnames(seurat)[seurat$CellType_II == cluster & seurat$Age == "P14"]
  avg <- rowMeans(as.matrix(seurat[["RNA"]]@data[gene_order,cells]))
  return(avg)
}
```

```{r}
receptor_expression_heatmap <- function(receptor_heatmap, ligand_heatmap,
                                      subset_name, plot_width, plot_height){
  
  #Get ligand list from heatmap
  gene_order <- receptor_heatmap$gtable$grobs[[2]]$label
  clusters <- ligand_heatmap$gtable$grobs[[2]]$label
  
  receptor_avg <- lapply(clusters, getAvgExp, gene_order=gene_order)
  receptor_avg <- do.call("cbind", receptor_avg)
  colnames(receptor_avg) <- clusters
  
  receptor_avg <- t(receptor_avg)
  
  ligand_exp_heatmap <- pheatmap(receptor_avg,
         color = colorRampPalette(c("whitesmoke", "purple"))(100),
         breaks = seq(0, 2, length.out=101),
         cluster_rows=F,
         cluster_cols=F,
         show_rownames=T,
         show_colnames=T,
         treeheight_row=0,
         treeheight_col=0,
         legend=F,
         border_color = "black",
         filename=paste0("C:/Users/.../", subset_name, "_receptor_expression_NOX.png"),
         width=plot_width, 
         height=plot_height)
}
```

```{r}
receptor_expression_heatmap(receptor_stroma, ligand_stroma, "Stroma", 17, 1.8)
receptor_expression_heatmap(receptor_myeloid, ligand_myeloid, "Myeloid", 20, 1.6)
receptor_expression_heatmap(receptor_endothelial, ligand_endothelial,"Endothelial", 12, 1.3)
receptor_expression_heatmap(receptor_epithelial, ligand_epithelial, "Epithelial", 17, 1.5)
receptor_expression_heatmap(receptor_lymphocytes, ligand_lymphocytes,"Lymphocytes", 12, 1.1)
receptor_expression_heatmap(receptor_mesothelial, ligand_mesothelial, "Mesothelial", 8, 0.85)
```





# Circos plot
```{r}
install.packages("DMwR2")
install.packages("Polychrome")
install.packages("circlize")
```

```{r}
library(DMwR2)
library(Polychrome)
library(circlize)
```


# Turn on output device
```{r}
pdf("C:/Users/.../Circos_plot.pdf",
    width=10, height=10)
```

# Load data in
# Organize by cell type groups -> only write ligand senders here
```{r}
cell_type_order <- c( #Epithelial
                     "Art", "gCap", "Vein", #Endothelial
                     "Myofibroblast", "Ly6a+ L-MSC", "Pericyte2", #Stromal
                     "AT2", 
                     "DC1", #Most immune
                     "Int Mf")
```

```{r}
test_dat <- read.csv("C:/Users/../LR_summary_NOX.csv")
test_dat <- arrange(test_dat, match(Sender, cell_type_order), Ligand)
test_dat$from <- paste0(test_dat$Sender, "_", test_dat$Ligand)
test_dat$to <- test_dat$Receiver
test_dat$value <- 1
df <- data.frame(from = test_dat$from,
                 to = test_dat$to, 
                 value= test_dat$value)
```


#Set up break distances between sectors -> these are sectors for ligand senders
```{r}
ligand_counts <- test_dat[,c("Sender", "Ligand")]
ligand_counts <- unique(ligand_counts)
ligand_counts <- table(ligand_counts$Sender)[cell_type_order]
receiver_count <- length(unique(test_dat$Receiver))

gaps <- c(rep(1.25, ligand_counts[1] - 1), 5,
          rep(1.25, ligand_counts[2] - 1), 5,
          rep(1.25, ligand_counts[3] - 1), 5,
          rep(1.25, ligand_counts[4] - 1), 5,
          rep(1.25, ligand_counts[5] - 1), 5,
          rep(1.25, ligand_counts[6] - 1), 5,
          rep(1.25, ligand_counts[7] - 1), 5,
          rep(1.25, ligand_counts[8] - 1), 5,
          rep(1.25, ligand_counts[9] - 1), 10,#Left sender-receiver border
          rep(1.25, receiver_count-1), 10) #receiver end
```


#Sender cell types
```{r}
sender_list <- unique(test_dat$Sender)
receiver_list <- unique(test_dat$Receiver)
```

# Prepare colours
```{r}
pdf("C:/Users/.../Circos_plot.pdf",
    width=10, height=10)


cols <- kelly.colors()[2:18]
names(cols) <- sender_list
test_dat$cols <- cols[test_dat$Sender] 
cols <- test_dat[,c("from", "cols")]
cols <- unique(cols)
tmp_names <- cols$from
cols <- cols$cols
names(cols) <- tmp_names
cols <- c(cols, rep("grey40", length(receiver_list)))
names(cols)[1] <- receiver_list #Adjust index to appropriate length

circos.clear()
circos.par(track.height = 0.125, gap.after = gaps)

chordDiagram(df,annotationTrack = c("grid"), scale=F,
             preAllocateTracks = 2, 
             directional = 1, 
             grid.col = cols
             #direction.type = "arrows",
             #link.arr.width = 0.15,
             #link.arr.length = 0.15,
             #big.gap = 20, small.gap = 1
             )

tmp <- test_dat[,c("from", "Ligand")]
tmp <- unique(tmp)
###

circos.track(track.index = 2, track.height = uh(1, "inches"),
             panel.fun = function(x, y) {
               circos.text(CELL_META$xcenter, CELL_META$ylim[1], tmp$Ligand[CELL_META$sector.numeric.index], 
                           facing = "clockwise", cex = 0.75,
                           niceFacing = TRUE, adj = c(0, 0.5))}, 
             bg.border = NA) # here set bg.border to NA is important


for(i in 1:length(sender_list)){
  ligand_list <- test_dat %>% filter(Sender == sender_list[i]) %>% pull(from)
  highlight.sector(ligand_list, track.index = 1, col = "grey80",
                   text = "", cex = 0.8, text.col = "grey80",
                   niceFacing = T, border = T, padding = c(0,0, -0.75, 0))
}

dev.off()
```


















