---
title: "R Notebook"
output: html_notebook
---

```{r}
library(Seurat)
library(ggplot2)
library(viridis)
library(dplyr)
library(pheatmap)
library(tidyr)
```

```{r}
MSCs <- readRDS("C:/Users/.../seurat_filtered.rds")
```

```{r}
DefaultAssay(MSCs) <- "RNA"
```


#Add Ly6a status
```{r}
MSCs$Ly6a_status <- "NA"
MSCs$Ly6a_status[MSCs$Sample %in% c("HYX1+", "HYX2+", "HYX3+", "NOX1+", "NOX2+", "NOX3+")] <- "Ly6a+"
MSCs$Ly6a_status[MSCs$Sample %in% c("HYX1-", "HYX2-", "HYX3-", "NOX1-", "NOX2-", "NOX3-")] <- "Ly6a-"
```
```{r}
MSCs$Ly6a_status <- factor(MSCs$Ly6a_status)
```

#Add Groups
```{r}
MSCs$Groups <- "NA"
MSCs$Groups[MSCs$Sample %in% c("HYX1+", "HYX2+", "HYX3+")] <- "Ly6a+_HYX"
MSCs$Groups[MSCs$Sample %in% c("HYX1-", "HYX2-", "HYX3-")] <- "Ly6a-_HYX"
MSCs$Groups[MSCs$Sample %in% c("NOX1+", "NOX2+", "NOX3+")] <- "Ly6a+_NOX"
MSCs$Groups[MSCs$Sample %in% c("NOX1-", "NOX2-", "NOX3-")] <- "Ly6a-_NOX"
```
```{r}
MSCs$Groups <- factor(MSCs$Groups)
```

#Save
```{r}
saveRDS(MSCs, file="C:/Users/.../seurat_filtered.rds")
```



#Find cluster markers
```{r}
markers <- FindAllMarkers(MSCs, logfc.threshold=1,only.pos=T)
```
```{r}
write.csv(markers, file="C:/Users/.../MSCs_Cluster markers.csv", quote=F)
```



#MAKE SIMPLIFIED HEATMAP
```{r}
MSC.markers <- read.csv("C:/Users/.../MSCs_Cluster markers.csv",row.names=1)
```
# Get list of clusters/cell types
```{r}
cluster_list <- levels(Idents(MSCs))
```
# Get top markers
```{r}
top_markers <- MSC.markers %>% 
  group_by(cluster) %>% 
  top_n(n=10, wt = avg_log2FC) %>% 
  pull(gene) 
```

```{r}
getAvgExp <- function(cluster){
  cells <- colnames(MSCs)[Idents(MSCs) == cluster] 
  avg <- rowMeans(as.matrix(MSCs[["RNA"]]@data[top_markers, cells])) 
  return(avg)}
```
```{r}
avg_exp <- lapply(cluster_list, getAvgExp)
avg_exp <- do.call("cbind", avg_exp)
colnames(avg_exp) <- cluster_list
```
```{r}
avg_exp <- t(scale(t(avg_exp), scale=T, center=T))
```
```{r}
pheatmap(t(avg_exp),
         color = viridis::inferno(105),
         cluster_rows = F,
         cluster_cols = F,
         fontsize = 15,
         border_color="black", #color around heatmap cells. Change as needed
         breaks = seq(-0.9, 2, length.out=101), # this is the range of values for the colormap, adjust as desired. Usually has to be modified to get good contrast
         legend=T, #switch to FALSE if the legend is getting in the way
         filename = "C:/Users/.../MSCs_heat map.png", dpi=600, width =12, height = 2.7) #adjust width/height to make dimensions work
```



##Dotplot from all clusters
```{r}
Idents(MSCs) <- "Groups"
```

#Differential expression comparisons
```{r}
Diff_exp_Ly6a_status_NOX <- FindMarkers(MSCs, ident.1="Ly6a+_NOX", ident.2="Ly6a-_NOX", group.by="Groups")
```
```{r}
write.csv(Diff_exp_Ly6a_status_NOX, file="C:/Users/.../Diff_exp_Ly6a_status_NOX.csv", quote=F)
```

```{r}
Diff_exp_Ly6a_status_HYX <- FindMarkers(MSCs, ident.1="Ly6a+_HYX", ident.2="Ly6a-_HYX", group.by="Groups")
```
```{r}
write.csv(Diff_exp_Ly6a_status_HYX, file="C:/Users/mizik/.../Diff_exp_Ly6a_status_HYX.csv", quote=F)
```

```{r}
Diff_exp_Ly6a_neg_changes_HYX <- FindMarkers(MSCs, ident.1="Ly6a-_HYX", ident.2="Ly6a-_NOX", group.by="Groups")
```
```{r}
write.csv(Diff_exp_Ly6a_neg_changes_HYX, file="C:/Users/mizik/.../Diff_exp_Ly6a_neg_changes_HYX.csv", quote=F)
```

```{r}
Diff_exp_Ly6a_pos_changes_HYX <- FindMarkers(MSCs, ident.1="Ly6a+_HYX", ident.2="Ly6a+_NOX", group.by="Groups")
```
```{r}
write.csv(Diff_exp_Ly6a_pos_changes_HYX, file="C:/Users/.../Diff_exp_Ly6a_pos_chnages_HYX.csv", quote=F)
```



#Find cluster markers
```{r}
Idents(MSCs)<- "Groups"
markers <- FindAllMarkers(MSCs, logfc.threshold=0.5,only.pos=T)
```
```{r}
write.csv(markers, file="C:/Users/.../MSCs_Groups markers.csv", quote=F)
```


# Cell frequency
```{r}
clusters_freq <- MSCs@meta.data %>%
  group_by(Sample, seurat_clusters) %>%
  summarise(n=n()) %>%
  mutate(relative_freq = n/sum(n))
```
```{r}
write.csv(clusters_freq, file="C:/Users/mizik/.../MSCs_clusters_cell_frequency.csv", quote=F,
          row.names = F)
```


#Plot most promising new markers

#MAKE SIMPLIFIED HEATMAP_best markers
```{r}
Idents(MSCs)<- "Groups"
```

```{r}
MSCs.markers <- read.csv("C:/Users/.../best markers_P7 MSCs.csv",row.names=1)
```
# Get list of clusters/cell types
```{r}
cluster_list <- levels(Idents(MSCs))
```
# Get top markers
```{r}
top_markers <- MSCs.markers %>% 
  group_by(cluster) %>% 
  pull(gene) 
```
# Make function to calculate average expression in each cluster
```{r}
getAvgExp <- function(cluster){
  cells <- colnames(MSCs)[Idents(MSCs) == cluster] 
  avg <- rowMeans(as.matrix(MSCs[["RNA"]]@data[top_markers, cells])) 
  return(avg)}
```
# Calculate average expression of top_markers in each cluster
```{r}
avg_exp <- lapply(cluster_list, getAvgExp)
avg_exp <- do.call("cbind", avg_exp)
colnames(avg_exp) <- cluster_list
```
# Transform values to Z-scores
```{r}
avg_exp <- t(scale(t(avg_exp), scale=T, center=T))
```
# Make heatmap
```{r}
pheatmap(t(avg_exp),
         color = viridis::inferno(105),
         cluster_rows = F, 
         cluster_cols = F,
         border_color="black", #color around heatmap cells. Change as needed
         breaks = seq(-0.5, 2, length.out=101), # this is the range of values for the colormap, adjust as desired. Usually has to be modified to get good contrast
         legend=T, #switch to FALSE if the legend is getting in the way
         filename = "C:/Users/.../Simplified heatmap_in vitro_markers_P7.png", dpi=600, width = 15, height = 2.3) #adjust width/height to make dimensions work
```



#Dotplot_Best potential L-MSC markers
```{r}
df <- as.matrix(MSCs[["RNA"]]@data[c("Acta2", "Actg2", "Adamts5", "C3", "Cemip",  "Col1a1", "Col1a2", "Col3a1", "Cygb", "Fxyd5", "Hmga2", "Lum", "Lxn", "Meg3", "Mfap5", "Nrn1", "Postn", "Prdx6", "Ptn", "Rbp1", "Serpina3n", "Serpinf1", "Tagln2", "Timp1", "Tm4sf1", "Wisp2", "Dcn"),])

#Subset for only Cap and Cap-a
df <- df[,MSCs$Groups %in% c("Ly6a-_HYX", "Ly6a+_HYX", "Ly6a-_NOX", "Ly6a+_NOX")]

df <- scale(t(df), scale=T, center=T)
df <- as.data.frame(df)

#Get metadata from seurat
df$Groups <- paste0(MSCs$Groups[rownames(df)], " ", MSCs$Oxygen[rownames(df)])

#Tidy
df <- gather(df, key="Gene", value="Expression", -Groups)

#Summarize expression
df <- df %>%
  group_by(Groups, Gene) %>%
  summarize(AvgExp = mean(Expression), Frequency = (sum(Expression > 0) / length(Expression))*100)
```

Plot
```{r}
dot_plot <- ggplot(df, aes(x=reorder(Gene, AvgExp), y=Groups)) +
  geom_point(color="black", pch=21, aes(size=Frequency, fill=AvgExp)) +
  scale_size(range=c(0,15), name = "Percent Expressed") +
  scale_fill_viridis(option="C", name="Relative Expression") +
  theme_classic() +
  theme(axis.text.x = element_text(angle=45, hjust=1, color="black", size=15),
        axis.text.y = element_text(size=17, color="black"),
        axis.title = element_blank())

dot_plot
```

```{r}
ggsave(dot_plot, filename = "C:/Users/.../Best L-MSC markers_in vitro.png", width=20, height=4.6, dpi=600)
```

