---
title: "R Notebook"
output: html_notebook
---

```{r}
library(Seurat)
library(ggplot2)
library(viridis)
library(dplyr)
library(pheatmap)
library(tidyr)
```

```{r}
memory.limit(size=90000)
stromal <- readRDS("C:/Users/.../stromal_cells.rds")
```

```{r}
stromal = UpdateSeuratObject(object = stromal)
```
```{r}
DefaultAssay(stromal) <- "RNA"
```


#Add CellType
```{r}
stromal$CellType <- "NA"
stromal$CellType[stromal$seurat_clusters %in% c("0")] <- "Col13a1+ fibroblast"
stromal$CellType[stromal$seurat_clusters %in% c("1")] <- "Myofibroblast"
stromal$CellType[stromal$seurat_clusters %in% c("2")] <- "Col14a1+ fibroblast"
stromal$CellType[stromal$seurat_clusters %in% c("3")] <- "Pericyte 1"
stromal$CellType[stromal$seurat_clusters %in% c("4")] <- "Pericyte 2"
stromal$CellType[stromal$seurat_clusters %in% c("5")] <- "Fibromyo/SMCs"
```
```{r}
stromal$CellType <- factor(stromal$CellType)
levels(stromal$CellType)
```


#Add Ly6a Groups
```{r}
stromal$Groups<- "NA"
stromal$Groups[stromal[["RNA"]]@data["Ly6a",] > 0] <- "Ly6a+"
stromal$Groups[stromal[["RNA"]]@data["Ly6a",]  == 0] <- "Ly6a-"
```
```{r}
stromal$Groups <- factor(stromal$Groups)
levels(stromal$Groups)
```


#Add Ly6a+ L-MSC (CellType_II)
```{r}
stromal$CellType_II <- "NA"
stromal$CellType_II[stromal$Groups== "Ly6a+"] <- "Ly6a+ L-MSC"
stromal$CellType_II[stromal$Groups== "Ly6a-" & stromal$CellType== "Col13a1+ fibroblast"] <- "Col13a1+ fibroblast"
stromal$CellType_II[stromal$Groups== "Ly6a-" & stromal$CellType== "Myofibroblast"] <- "Myofibroblast"
stromal$CellType_II[stromal$Groups== "Ly6a-" & stromal$CellType== "Col14a1+ fibroblast"] <- "Col14a1+ fibroblast"
stromal$CellType_II[stromal$Groups== "Ly6a-" & stromal$CellType== "Pericyte 1"] <- "Pericyte 1"
stromal$CellType_II[stromal$Groups== "Ly6a-" & stromal$CellType== "Pericyte 2"] <- "Pericyte 2"
stromal$CellType_II[stromal$Groups== "Ly6a-" & stromal$CellType== "Fibromyo/SMCs"] <- "Fibromyo/SMCs"
```
```{r}
stromal$CellType_II <- factor(stromal$CellType_II)
levels(stromal$CellType_II)
```

#Add Ly6aGroups
```{r}
stromal$Ly6aGroups <- "NA"
stromal$Ly6aGroups[stromal$Groups %in% c("Ly6a+") & stromal$Oxygen %in% c("Normoxia")] <- "Ly6a+_NOX"
stromal$Ly6aGroups[stromal$Groups %in% c("Ly6a+") & stromal$Oxygen %in% c("Hyperoxia")] <- "Ly6a+_HYX"
stromal$Ly6aGroups[stromal$Groups %in% c("Ly6a-") & stromal$Oxygen %in% c("Normoxia")] <- "Ly6a-_NOX"
stromal$Ly6aGroups[stromal$Groups %in% c("Ly6a-") & stromal$Oxygen %in% c("Hyperoxia")] <- "Ly6a-_HYX"
```
```{r}
stromal$Ly6aGroups <- factor(stromal$Ly6aGroups)
levels(stromal$Ly6aGroups)
```


#Save point
```{r}
saveRDS(stromal, file="C:/Users/.../stromal_cells.rds")
```


#Find markers for CellType_II
```{r}
Idents(stromal)<- "CellType_II"
```
```{r}
markers <- FindAllMarkers(stromal, logfc.threshold=1, only.pos=T)
write.csv(markers, file="C:/Users/.../stroma_cluster_markers.csv", quote=F)
```

```{r}
stromal.markers <- read.csv("C:/Users/.../stroma_cluster_markers.csv",row.names=1)
```
```{r}
stromal.markers_II %>% group_by(cluster) %>% top_n(10, wt=avg_log2FC) 
```


#MAKE SIMPLIFIED HEATMAP
```{r}
Idents(stromal)<- "CellType_II"
```

```{r}
stromal.markers <- read.csv("C:/Users/.../stroma_cluster_markers_top 10.csv",row.names=1)
```
# Get list of clusters/cell types
```{r}
cluster_list <- levels(Idents(stromal))
```
# Get top markers
```{r}
top_markers <- stromal.markers %>% 
  group_by(cluster) %>% 
  top_n(n=10, wt = avg_log2FC) %>% 
  pull(gene) 
```
# Make function to calculate average expression in each cluster
```{r}
getAvgExp <- function(cluster){
  cells <- colnames(stromal)[Idents(stromal) == cluster] 
  avg <- rowMeans(as.matrix(stromal[["RNA"]]@data[top_markers, cells])) 
  return(avg)}
```
# Calculate average expression of top_markers in each cluster
```{r}
avg_exp <- lapply(cluster_list, getAvgExp)
avg_exp <- do.call("cbind", avg_exp)
colnames(avg_exp) <- cluster_list
```
# Transform values to Z-scores
```{r}
avg_exp <- t(scale(t(avg_exp), scale=T, center=T))
```
# Make heatmap
```{r}
pheatmap(t(avg_exp),
         color = viridis::inferno(105),
         cluster_rows = F,
         cluster_cols = F, 
         border_color="black", 
         breaks = seq(-0.5, 2, length.out=101), 
         legend=T, 
         filename = "C:/Users/.../Stroma_cluster markers_top10.png", dpi=600, width = 12, height = 2.5) 
```



# Calculate the number of cells in each Ly6aGroup for each individual animal and time point
```{r}
stromal$UniqueSample <- paste(stromal$orig.ident, stromal$Oxygen, stromal$Barcode, sep = "_")

clusters_freq <- stromal@meta.data %>%
  group_by(UniqueSample, Ly6aGroups) %>%
  summarise(n=n()) %>%
  mutate(relative_freq = n/sum(n))

write.csv(clusters_freq, file="C:/Users/.../Ly6aGroups_cell_frequency.csv", quote=F,
          row.names = F)
```


#Subset NOX cells
```{r}
cells_keep <- colnames(stromal)[stromal$Oxygen %in% c("Normoxia")]
stromal_NOX <- subset(stromal, cells=cells_keep)
saveRDS(stromal_NOX, file="C:/Users/.../stromal_cells_NOX.rds")
```
```{r}
stromal_NOX <- readRDS("C:/Users/.../stromal_cells_NOX.rds")
```

#DotPlot_ NOX _gene expression in Ly6a L-MSC
```{r}
df <- as.matrix(stromal_NOX[["RNA"]]@data[c("Mmp3", "Inmt", "Aldh1a1", "Hlf", "Mettl7a1", "Enpp2", "Podn", "Abca1", "Apoe", "Scube2", "Cxcl14", "Per3", "Pdzd2", "Epas1", "Tmcc3", "Klf9", "Foxo1", "Gdf10", "Mt1", "H2-K1","Top2a", "Cenpf", "Pclaf", "Ube2c", "Cks2", "Cdk1", "Mest", "Ptn", "H19", "Lmnb1", "Agtr2", "H2afx", "Smc2", "Fbn2", "Dlk1", "Hmgn2", "Stmn1", "Hmgb2", "Nrk", "Dut"),])

df <- df[,stromal_NOX$CellType_II %in% c("Ly6a+ L-MSC")]

df <- scale(t(df), scale=T, center=T)
df <- as.data.frame(df)

df$CellType_II <- paste0(stromal_NOX$CellType_II[rownames(df)], " ", stromal_NOX$Age[row.names(df)])

df <- gather(df, key="Gene", value="Expression", -CellType_II)

df <- df %>%
  group_by(CellType_II, Gene) %>%
  summarize(AvgExp = mean(Expression), Frequency = (sum(Expression > 0) / length(Expression))*100)
```

```{r}
write.csv(df, file="C:/Users/.../df_CellTypeII.csv", quote=F)
```
```{r}
df <- read.csv("C:/Users/.../df_CellTypeII.csv",row.names=1)
```

Plot
```{r}
dot_plot <- ggplot(df, aes (x=reorder(Gene, AvgExp), y= CellType_II)) +
  geom_point(color="black", pch=21, aes(size=Frequency, fill=AvgExp)) +
  scale_size(range=c(0,15), name = "Percent Expressed", limits = c(0, 100)) +
  scale_fill_viridis(option="D", name="Relative Expression") +
  theme_classic() +
  theme(axis.text.x = element_text(angle=45, hjust=1, color="black", size=15),
        axis.text.y = element_text(size=15, color="black"),
        axis.title = element_blank()) 
dot_plot
```

```{r}
ggsave(dot_plot, filename = "C:/Users/.../GellTypeII genes_NOX.png", width=21, height=4.75, dpi=600)
```


#DotPlot_NOX_only genes of Ly6a+ L-MSCs
```{r}
df <- as.matrix(stromal_NOX[["RNA"]]@data[c("Mmp3", "Podn", "Mt1", "C1s1", "Timp3", "Igfbp6", "Abca8a", "C3", "Agtr2", "Dlk1", "Cdkn1c"),])

df <- df[,stromal_NOX$Groups %in% c("Ly6a+", "Ly6a-")]
df <- scale(t(df), scale=T, center=T)
df <- as.data.frame(df)

df$Groups <- paste0(stromal_NOX$Groups[rownames(df)], " ", stromal_NOX$Age[rownames(df)])

df <- gather(df, key="Gene", value="Expression", -Groups)

df <- df %>%
  group_by(Groups, Gene) %>%
  summarize(AvgExp = mean(Expression), Frequency = (sum(Expression > 0) / length(Expression))*100)
```

```{r}
write.csv(df, file="C:/Users/../df_groups.csv", quote=F)
```

```{r}
df <- read.csv("C:/Users/.../df_groups.csv",row.names=1)
```

Plot
```{r}
dot_plot <- ggplot(df, aes(x=reorder(Gene, Order),y=Groups)) +
  geom_point(color="black", pch=21, aes(size=Frequency, fill=AvgExp)) +
  scale_size (range=c(0,15), name = "Percent Expressed", limits = c(0,100)) +
  scale_fill_viridis(option="D", name="Relative Expression") +
  theme_classic() +
  theme(axis.text.x = element_text(angle=45, hjust=1, color="black", size=20),
        axis.text.y = element_text(size=20, color="black"),
        axis.title = element_blank())
dot_plot
```
```{r}
ggsave(dot_plot, filename = "C:/Users/.../Groups genes_NOX.png", width=10, height=5, dpi=600)
```





#Plot most promising new markers in P7 cells
```{r}
DefaultAssay(stromal) <- "RNA"
stromal_P7 <- colnames(stromal)[stromal$Age %in% c("P7")]
stromal_P7<- subset(stromal, cells=stromal_P7)
```


#MAKE SIMPLIFIED HEATMAP_best markers
```{r}
Idents(stromal_P7)<- "Ly6aGroups"
```

```{r}
stromal.markers <- read.csv("C:/Users/.../best markers_P7.csv",row.names=1)
```
# Get list of clusters/cell types
```{r}
cluster_list <- levels(Idents(stromal_P7))
```
# Get top markers
```{r}
top_markers <- stromal.markers %>% 
  group_by(cluster) %>% 
  pull(gene) 
```
# Make function to calculate average expression in each cluster
```{r}
getAvgExp <- function(cluster){
  cells <- colnames(stromal_P7)[Idents(stromal_P7) == cluster] 
  avg <- rowMeans(as.matrix(stromal_P7[["RNA"]]@data[top_markers, cells])) 
  return(avg)}
```
# Calculate average expression of top_markers in each cluster
```{r}
avg_exp <- lapply(cluster_list, getAvgExp)
avg_exp <- do.call("cbind", avg_exp)
colnames(avg_exp) <- cluster_list
```
# Transform values to Z-scores
```{r}
avg_exp <- t(scale(t(avg_exp), scale=T, center=T))
```
# Make heatmap
```{r}
pheatmap(t(avg_exp),
         color = viridis::inferno(105),
         cluster_rows = F, 
         cluster_cols = F,
         border_color="black",
         breaks = seq(-0.5, 2, length.out=101), 
         legend=T,
         filename = "C:/Users/.../Simplified heatmap_in vivo_markers_P7.png", dpi=600, width = 15, height = 2.3)
```


#Dotplot_Best potential L-MSC markers
```{r}
df <- as.matrix(stromal_P7[["RNA"]]@data[c("Acta2", "Actg2", "Col14a1", "Col1a1", "Col1a2", "Col3a1", "Clec3b", "Cygb", "Dcn", "Fst", "Itm2a", "Igfbp6", "Ifi27l2a", "Has1", "Il33", "Lum", "Meg3", "Mfap5", "Nrk", "Postn", "Pi16", "Prrx1", "Serpina3n", "Serpinf1", "Timp1", "Tnxb", "S100a10", "S100a6", "Ptn", "Cxcl1", "Akap12", "C3"),])

df <- df[,stromal_P7$Groups %in% c("Ly6a-", "Ly6a+")]

df <- scale(t(df), scale=T, center=T)
df <- as.data.frame(df)

df$Groups <- paste0(stromal_P7$Groups[rownames(df)], " ", stromal_P7$Oxygen[rownames(df)])

df <- gather(df, key="Gene", value="Expression", -Groups)

df <- df %>%
  group_by(Groups, Gene) %>%
  summarize(AvgExp = mean(Expression), Frequency = (sum(Expression > 0) / length(Expression))*100)
```

Plot
```{r}
dot_plot <- ggplot(df, aes(x=reorder(Gene, AvgExp), y=Groups)) +
  geom_point(color="black", pch=21, aes(size=Frequency, fill=AvgExp)) +
  scale_size(range=c(0,15), name = "Percent Expressed") +
  scale_fill_viridis(option="C", name="Relative Expression") +
  theme_classic() +
  theme(axis.text.x = element_text(angle=45, hjust=1, color="black", size=15),
        axis.text.y = element_text(size=17, color="black"),
        axis.title = element_blank())

dot_plot
```

```{r}
ggsave(dot_plot, filename = "C:/Users/.../Best L-MSC markers_in vivo.png", width=20, height=4.6, dpi=600)
```
