---
title: "LT network_receiver_NOX"
output: html_notebook
---


```{r}
library(Seurat)
library(ggplot2)
library(viridis)
library(dplyr)
library(pheatmap)
library(nichenetr)
library(tidyverse)
library(maps)
```

#Read in the expression data of interacting cells:
```{r}
memory.limit(size=60000)
seuratObj = readRDS("C:/Users/.../seurat_lung_NOX.rds")
seuratObj@meta.data %>% head()
##         nGene nUMI orig.ident aggregate res.0.6 celltype nCount_RNA nFeature_RNA
## W380370   880 1611      LN_SS        SS       1    CD8 T       1607          876
## W380372   541  891      LN_SS        SS       0    CD4 T        885          536
## W380374   742 1229      LN_SS        SS       0    CD4 T       1223          737
## W380378   847 1546      LN_SS        SS       1    CD8 T       1537          838
## W380379   839 1606      LN_SS        SS       0    CD4 T       1603          836
## W380381   517  844      LN_SS        SS       0    CD4 T        840          513
```

```{r}
seuratObj$celltype <- seuratObj$CellType_II
```


#Read in NicheNet’s ligand-target prior model, ligand-receptor network and weighted integrated networks:
```{r}
ligand_target_matrix = readRDS(url("https://zenodo.org/record/3260758/files/ligand_target_matrix.rds"))
ligand_target_matrix[1:5,1:5] # target genes in rows, ligands in columns
##                 CXCL1        CXCL2        CXCL3        CXCL5         PPBP
## A1BG     3.534343e-04 4.041324e-04 3.729920e-04 3.080640e-04 2.628388e-04
## A1BG-AS1 1.650894e-04 1.509213e-04 1.583594e-04 1.317253e-04 1.231819e-04
## A1CF     5.787175e-04 4.596295e-04 3.895907e-04 3.293275e-04 3.211944e-04
## A2M      6.027058e-04 5.996617e-04 5.164365e-04 4.517236e-04 4.590521e-04
## A2M-AS1  8.898724e-05 8.243341e-05 7.484018e-05 4.912514e-05 5.120439e-05

lr_network = readRDS(url("https://zenodo.org/record/3260758/files/lr_network.rds"))
head(lr_network)
## # A tibble: 6 x 4
##   from  to    source         database
##   <chr> <chr> <chr>          <chr>   
## 1 CXCL1 CXCR2 kegg_cytokines kegg    
## 2 CXCL2 CXCR2 kegg_cytokines kegg    
## 3 CXCL3 CXCR2 kegg_cytokines kegg    
## 4 CXCL5 CXCR2 kegg_cytokines kegg    
## 5 PPBP  CXCR2 kegg_cytokines kegg    
## 6 CXCL6 CXCR2 kegg_cytokines kegg

weighted_networks = readRDS(url("https://zenodo.org/record/3260758/files/weighted_networks.rds"))
weighted_networks_lr = weighted_networks$lr_sig %>% inner_join(lr_network %>% distinct(from,to), by = c("from","to"))

head(weighted_networks$lr_sig) # interactions and their weights in the ligand-receptor + signaling network
## # A tibble: 6 x 3
##   from  to     weight
##   <chr> <chr>   <dbl>
## 1 A1BG  ABCC6  0.422 
## 2 A1BG  ACE2   0.101 
## 3 A1BG  ADAM10 0.0970
## 4 A1BG  AGO1   0.0525
## 5 A1BG  AKT1   0.0855
## 6 A1BG  ANXA7  0.457
head(weighted_networks$gr) # interactions and their weights in the gene regulatory network
## # A tibble: 6 x 3
##   from  to     weight
##   <chr> <chr>   <dbl>
## 1 A1BG  A2M    0.0294
## 2 AAAS  GFAP   0.0290
## 3 AADAC CYP3A4 0.0422
## 4 AADAC IRF8   0.0275
## 5 AATF  ATM    0.0330
## 6 AATF  ATR    0.0355
```


#Because the expression data is of mouse origin, we will convert the NicheNet network gene symbols from human to mouse based on one-to-one orthology:
```{r}
lr_network = lr_network %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
colnames(ligand_target_matrix) = ligand_target_matrix %>% colnames() %>% convert_human_to_mouse_symbols()
rownames(ligand_target_matrix) = ligand_target_matrix %>% rownames() %>% convert_human_to_mouse_symbols()

ligand_target_matrix = ligand_target_matrix %>% .[!is.na(rownames(ligand_target_matrix)), !is.na(colnames(ligand_target_matrix))]

weighted_networks_lr = weighted_networks_lr %>% mutate(from = convert_human_to_mouse_symbols(from), to = convert_human_to_mouse_symbols(to)) %>% drop_na()
```


#PERFORM THE NICHENET ANALYSIS:
```{r}
Idents(seuratObj) <- "celltype"
```


#ONLY CELLS BASED ON CIRCOS PLOT ARE SENDERS
```{r}
## receiver
receiver = c("Ly6a+ L-MSC")
expressed_genes_receiver = get_expressed_genes(receiver, seuratObj, assay_oi = "RNA", pct = 0.10)

background_expressed_genes = expressed_genes_receiver %>% .[. %in% rownames(ligand_target_matrix)]
```

#-> Changing the assay to RNA
```{r}
## sender
sender_celltypes = c("gCap", "Art", "Vein", "Int Mf", "DC1", "AT2",  "Myofibroblast", "Pericyte 2", "Ly6a+ L-MSC")

list_expressed_genes_sender = sender_celltypes %>% unique() %>% lapply(get_expressed_genes, seuratObj, assay_oi = "RNA", pct = 0.10) # lapply to get the expressed genes of every sender cell type separately here
expressed_genes_sender = list_expressed_genes_sender %>% unlist() %>% unique()
```

#2. Define a gene set of interest
```{r}
seurat_obj_receiver= subset(seuratObj, idents = receiver)
seurat_obj_receiver = SetIdent(seurat_obj_receiver, value = seurat_obj_receiver[["Age"]])

condition_oi = "P14"
condition_reference = "P3" 
  
DE_table_receiver = FindMarkers(object = seurat_obj_receiver, ident.1 = condition_oi, ident.2 = condition_reference, min.pct = 0.10) %>% rownames_to_column("gene")

geneset_oi = DE_table_receiver %>% filter(p_val_adj <= 0.05 & abs(avg_log2FC) >= 0.5) %>% pull(gene)
geneset_oi = geneset_oi %>% .[. %in% rownames(ligand_target_matrix)]
```

#3. Define a set of potential ligands
```{r}
ligands = lr_network %>% pull(from) %>% unique()
receptors = lr_network %>% pull(to) %>% unique()

expressed_ligands = intersect(ligands,expressed_genes_sender)
expressed_receptors = intersect(receptors,expressed_genes_receiver)

potential_ligands = lr_network %>% filter(from %in% expressed_ligands & to %in% expressed_receptors) %>% pull(from) %>% unique()
```

#4) Perform NicheNet ligand activity analysis
```{r}
ligand_activities = predict_ligand_activities(geneset = geneset_oi, background_expressed_genes = background_expressed_genes, ligand_target_matrix = ligand_target_matrix, potential_ligands = potential_ligands)

ligand_activities = ligand_activities %>% arrange(-pearson) %>% mutate(Rank = 1:nrow(ligand_activities))
ligand_activities
```
#Save the ligant activities table
```{r}
write.csv(ligand_activities, file="C:/Users/.../ligand activities table_NOX_receiver.csv", quote=F)
```

```{r}
ligand_activities <- read.csv("C:/Users/.../ligand activities table_NOX_receiver.csv",row.names=1)
```

#The different ligand activity measures
```{r}
best_upstream_ligands = ligand_activities %>% top_n(5, pearson) %>% arrange(-pearson) %>% pull(test_ligand) %>% unique()
```

#Plot the graph for quality control
```{r}
# show histogram of ligand activity scores
p_hist_lig_activity = ggplot(ligand_activities, aes(x=pearson)) + 
  geom_histogram(color="black", fill="darkorange")  + 
  # geom_density(alpha=.1, fill="orange") +
  geom_vline(aes(xintercept=min(ligand_activities %>% top_n(5, pearson) %>% pull(pearson))), color="red", linetype="dashed", size=1) + 
  labs(x="ligand activity (PCC)", y = "# ligands") +
  theme_classic()
p_hist_lig_activity
```

#5) Infer receptors and top-predicted target genes of ligands that are top-ranked in the ligand activity analysis

#Active target gene inference
```{r}
active_ligand_target_links_df = best_upstream_ligands %>% lapply(get_weighted_ligand_target_links,geneset = geneset_oi, ligand_target_matrix = ligand_target_matrix, n = 200) %>% bind_rows() %>% drop_na()

active_ligand_target_links = prepare_ligand_target_visualization(ligand_target_df = active_ligand_target_links_df, ligand_target_matrix = ligand_target_matrix, cutoff = 0.33)

order_ligands = intersect(best_upstream_ligands, colnames(active_ligand_target_links)) %>% rev() %>% make.names()
order_targets = active_ligand_target_links_df$target %>% unique() %>% intersect(rownames(active_ligand_target_links)) %>% make.names()
rownames(active_ligand_target_links) = rownames(active_ligand_target_links) %>% make.names() # make.names() for heatmap visualization of genes like H2-T23
colnames(active_ligand_target_links) = colnames(active_ligand_target_links) %>% make.names() # make.names() for heatmap visualization of genes like H2-T23

vis_ligand_target = active_ligand_target_links[order_targets,order_ligands] %>% t()
```

#-> Visualize:
```{r}
p_ligand_target_network = vis_ligand_target %>% make_heatmap_ggplot("Prioritized ligands","Predicted target genes", color = "purple",legend_position = "top", x_axis_position = "top",legend_title = "Regulatory potential")  + theme(axis.text.x = element_text(face = "italic")) + scale_fill_gradient2(low = "whitesmoke",  high = "purple", breaks = c(0,0.0045,0.0090))
p_ligand_target_network
```

#Save Heatmap
```{r}
p_ligand_target_network = vis_ligand_target %>% make_heatmap_ggplot("Prioritized ligands","Predicted target genes", color = "purple", legend_position = "top", x_axis_position = "top",legend_title = "Regulatory potential")   + scale_fill_gradient2(low = "gray",  high = "purple", mid= "gray95", breaks = c(0,0.0045,0.0090))

ggsave(p_ligand_target_network, filename="C:/Users/.../Targets/LT network_receiver_NOX.png", width=4, height=3, dpi=600)
```
